A cascaded supervised learning approach to inverse reinforcement learning

### 2012-11-04 piot2012cascading/ECML13_cascading.tex

% Created 2012-10-15 lun. 14:49
\documentclass[smallextended]{svjour3}
\smartqed 
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage[integrals]{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subfigure}
\tolerance=1000
%\usepackage{amsmath}
%\usepackage{amsthm}
%\newtheorem{definition}{Definition}
%\newtheorem{theorem}{Theorem}
\newtheorem{cor}{Corollary}
\usepackage{dsfont}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs=true}
\providecommand{\alert}[1]{\textbf{#1}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\newcommand{\Card}{\operatorname*{Card}} %\operatorname* pour les op. pouvant admettre des limites...
\begin{document}
\title{Learning a reward function from demonstrations: a cascaded supervised learning approach}
\author{Edouard Klein$^{1,2}$ \and Bilal Piot $^{2,3}$\and Matthieu Geist $^{2}$\and Olivier Pietquin$^{2,3}$}
\titlerunning{Cascading approach to IRL}
\institute{
 1. ABC Team\\
 LORIA-CNRS, France.\\
\and
2.Supélec\\
 IMS-MaLIS Research group, France\\
 \texttt{firstname.lastname@supelec.fr}\\
\and 
3. UMI 2958 GeorgiaTech-CNRS\\
France\\
}
\date{\today}

@@

\begin{abstract}
  This paper considers the Inverse Reinforcement Learning problem, that is inferring a reward function for which a demonstrated expert policy is optimal.
We propose to break the IRL problem down into two generic Supervised Learning steps: this is the Cascaded Supervised IRL (CSI) approach. A classification step that defines a score function is followed by a regression step that outputs a reward function.
An analysis of this approach shows that the demonstrated expert policy is near-optimal for the computed reward function.
Not needing to repeatedly solve a Markov Decision Process (MDP) and the ability to leverage existing techniques for classification and regression are two important advantages of the CSI approach. It is furthermore empirically demonstrated to work with only transitions sampled according to the expert policy, up to the use of some heuristics. For this purpose two classical benchmarks (a highway driving simulator and a inverted pendulum simulator) are used.
  \end{abstract}
\section{Introduction}
\label{sec-2}
The sequential decision making problem consists in taking the appropriate action pondering the available data in order to maximize a certain criterion. When framed in an MDP (see Sec.~\ref{sec:background}), the criterion to maximize is the expected sum of discounted rewards. The Inverse Reinforcement Learning (IRL)~\cite{russell1998learning} problem, which we tackle here, aims at inferring a reward function for which a demonstrated expert policy is optimal.

@@

This algorithm does not need to iteratively solve an MDP and needs only sampled transitions from expert and non-experts policies as inputs. Moreover, up to the use of some heuristics (see Subsubsec.~\ref{subsubsec:heuristics}), the algorithm is able to be trained only with transitions sampled from the demonstrated expert policy. A specific instantiation of CSI (proposed Subsec.~\ref{subsec:instanciation}) is tested on an inverted pendulum simulator (Subsec~\ref{subsec:IP}) and a car driving simulator (Subsec~\ref{subsec:Highway}) where we compare it with an AL algorithm \cite{abbeel2004apprenticeship}, with a pure classification algorithm \cite{taskar2005learning} and with a recent successfull IRL method \cite{klein2012scirl} as well as with a random baseline.

@@

We recall that solving the MDP is the direct problem. This contribution aims at solving the inverse one. We observe an expert's deterministic\footnotemark[\value{footnote}] policy $\pi_E$, assuming that there exists some unknown reward $R_E$ for which the expert is optimal. The suboptimality of the expert is an interesting setting that has been discussed for example in \cite{melo2010analysis,syed2010reduction}, but that we are not addressing here. We do not try to find this unknown reward $R_E$ but rather a non trivial reward $R$ for which the expert is at least near-optimal. The trivial reward $0$ is a solution to this ill-posed problem (no reward means that every behavior is optimal). Because of its ill-posed nature, this expression of \emph{Inverse Reinforcement Learning} (IRL) still has to find a satisfactory solution although lots of progress have been made, see Sec.~\ref{sec:related}.
\section{The cascading algorithm}
\label{sec:algo}
Our first step towards a reward function solving the IRL problem is a classification step using a \emph{score function-based multi-class classifier} (SFMC$^2$ for short). This classifier learns a score function $q\in\mathbb{R}^{S\times A}$ that rates the association of a given label $a\in A$ with a certain input $s\in S$. The classification rule $\pi_C\in A^S$ simply selects (one of) the label(s) that achieves the highest score for the given inputs:
\begin{equation}
  \label{eq:greedy2}
\pi_C(s) \in \argmax_a q(s,a).
\end{equation}
For example, \emph{Multi-class Support Vector Machines}~\cite{guermeur2011generic} can be seen as SFMC$^2$ algorithms, the same can be said of the structured margin approach \cite{taskar2005learning} which we consider in the experimental setting. Other algorithms may be envisionned (see Subsec.~\ref{subsec:instanciation}).

@@

The only constraints the bound of Thm.~\ref{thm} implies on datasets $D_R$ and $D_C$ is that they provide enough information to the supervised algorithms to keep both error terms $\epsilon_C$ and $\epsilon_R$ low. In Sec.~\ref{sec:experiments} we deal with a dire lack of data in dataset $D_R$. We address the problem with the use of heuristics (Subsubsec.~\ref{subsubsec:heuristics}) in order to show the behavior of the CSI algorithm in somewhat realistic conditions.

@@

%% Feel good comment
\section{Experiments}
\label{sec:experiments}
In this section, we put our approach under stress to empirically demonstrate its behavior. We begin by providing information pertaining to both benchmarks: an explanation about the amount and source of the available data is given on Subsubsec.~\ref{subsubsec:data} and a precise description of the way we instanciated the CSI algorithm is given on subsubsections \ref{subsubsec:subroutines} and \ref{subsubsec:heuristics}.

We then use a classical RL benchmark, the Inverted Pendulum (Subsec.~\ref{subsec:IP}), to provide a qualitative illustration of the proposed CSI approach.

We finally provide quantitative results and comparison with state-of-the-art approaches on a Highway driving simulator (Subsec.~\ref{subsec:highway}).
\subsection{Instanciation}
\label{subsec:instanciation}
\subsubsection{Data scarcity}
\label{subsubsec:data}
In order to create a somewhat realistic environment, we decided that the CSI algorithm could neither use the transition probabilities nor a simulator. The only data it could access is sampled according to the expert policy. This corresponds for example to a situation where a costly system can only be controlled by a trained operator as a bad control sequence could lead to a system breakage.

@@

\subsection{Inverted Pendulum}
\label{subsec:IP}
We provide here qualitative results on the inverted pendulum problem. Quantitative results obtained by following a procedure very similar to Subsec.~\ref{subsec:highway} were of the same nature and justified the same analysis as those provided in Subsubsec.~\ref{subsubsec:hresults}. We thus omit them here for the sake of avoiding redundancy. Instead, we use the inverted pendulum problem for illustrative purposes.
\subsubsection{Setting}
The inverted pendulum (also known as the cart pole) is a classical benchmark in the RL community, it is fully described in \cite{lagoudakis2003least}. The goal is to maintain a pole in a unstable equilibrium position with the help of a cart moving along one axis. The state space consists in the angular position an speed of the pole. It is possible to use an RL algorithm to learn how to balance the pendule in an upright position by using the following reward: as long as the angular position of the pole is (in absolute value) less than $\pi \over 2$ the reward is $0$, the reward is $-1$ if the pole falls ; this reward function is plotted on Fig.~\ref{toto}. Three actions are available to the agent: it can choose to apply a leftward or rightward force on the cart or no force at all.

@@

\subsection{Highway driving simulator}
\label{subsec:highway}
Let's now move from a qualitative illustration to a quantitative assessment of the performance of the CSI algorithm. The Highway driving problem is described Subsubsec.~\ref{subsubsec:hsetting}, the criterion we use is explained Subsubsec.~\ref{subsubsec:hcriterion} and finally results and comparison to other approaches are presented Subsubsec.~\ref{subsubsec:hresults}.
\subsubsection{Setting}
\label{subsubsec:hsetting}
The setting of the experiment is a driving simulator inspired from a benchmark already used in \cite{syed2008apprenticeship,syed2008game}. The agent controls a car that can switch between the three lanes of the road, go off-road on either side and modulate between three speed levels. At all times, there will be one car in one of the three lanes. Even at the lowest speed, the player's car moves faster than the others. When the other car disappears at the bottom of the screen, another one appear at the top in a randomly chosen lane. It takes two transitions to completely change lanes, as the player can move left or right for half a lane's length at a time. At the highest speed setting, if the other car appear in the lane the player is in, it is not possible to avoid the collision. The main difference between the original benchmark \cite{syed2008apprenticeship,syed2008game} and ours is that we made the MDP more ergodic by allowing the player to change speed whenever he wishes so, not just during the first transition. If anything, by adding two actions, we enlarged the state-action space and thus made the problem harder.
%FIXME:METTRE UNE CAPTURE DECRAN.
The reward function $R_E$ trains the expert to go as fast as possible (high reward) while avoiding collisions (harshly penalised) and avoiding going off-road (moderately penalised). Any other situation receives a null reward.

@@

  \caption{Data is shown with mean, standard deviation, minimum and maximum value over 50 runs. See Subsec.~\ref{subsec:setting} for detailled information about the experimental setting.}
  \label{fig:Highway}
\end{figure}

@@

Classification and IRL have met in the past in \cite{ratliff2006maximum}, but the labels were complete optimal policies rather than actions and the inputs were MDPs, which had to be solved. The SCIRL (for Structured Classification for IRL) algorithm \cite{klein2012scirl} which we used Subsec.~\ref{subsec:highway} is able to solve the IRL problem (without having to solve any MDP) in the same dire conditions the cascading approach has been tested in. It injects the structure of the MDP in a linearly parametrized score-function based classifier by using the expert's feature expectation as feature function. This necessity of a linarly parametrized approach and the need to compute the expert's feature exepctations makes the SCIRL algorithm less convenient than the CSI approach.

>>>>

% Created 2012-10-15 lun. 14:49
\documentclass[smallextended]{svjour3}
\smartqed 
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage[integrals]{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subfigure}
\tolerance=1000
%\usepackage{amsmath}
%\usepackage{amsthm}
%\newtheorem{definition}{Definition}
%\newtheorem{theorem}{Theorem}
\newtheorem{cor}{Corollary}
\usepackage{dsfont}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs=true}
\providecommand{\alert}[1]{\textbf{#1}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\newcommand{\Card}{\operatorname*{Card}} %\operatorname* pour les op. pouvant admettre des limites...
\begin{document}
\title{A cascaded supervised learning approach to inverse reinforcement learning}
\author{Edouard Klein$^{1,2}$ \and Bilal Piot $^{2,3}$\and Matthieu Geist $^{2}$\and Olivier Pietquin$^{2,3}$}
\titlerunning{Cascading approach to IRL}
\institute{
 1. ABC Team\\
 LORIA-CNRS, France.\\
\and
2.Supélec\\
 IMS-MaLIS Research group, France\\
 \texttt{firstname.lastname@supelec.fr}\\
\and 
3. UMI 2958 GeorgiaTech-CNRS\\
France\\
}
\date{\today}

@@

\begin{abstract}
  This paper considers the Inverse Reinforcement Learning (IRL) problem, that is inferring a reward function for which a demonstrated expert policy is optimal.
We propose to break the IRL problem down into two generic Supervised Learning steps: this is the Cascaded Supervised IRL (CSI) approach. A classification step that defines a score function is followed by a regression step providing a reward function.
A theoretical analysis shows that the demonstrated expert policy is near-optimal for the computed reward function.
Not needing to repeatedly solve a Markov Decision Process (MDP) and the ability to leverage existing techniques for classification and regression are two important advantages of the CSI approach. It is furthermore empirically demonstrated to work with only transitions sampled according to the expert policy, up to the use of some heuristics. This is exemplified on two classical benchmarks (a highway driving simulator and an inverted pendulum simulator).
  \end{abstract}
\section{Introduction}
\label{sec-2}
Sequential decision making consists in choosing the appropriate action given the available data in order to maximize a certain criterion. When framed in a Markov Decision Process (MDP) (see Sec.~\ref{sec:background}), the criterion to maximize is the expected sum of discounted rewards. The Inverse Reinforcement Learning (IRL)~\cite{russell1998learning} problem, which we tackle here, aims at inferring a reward function for which a demonstrated expert policy is optimal.

@@

This algorithm does not need to iteratively solve an MDP and needs only sampled transitions from expert and non-experts policies as inputs. Moreover, up to the use of some heuristics (see Sec.~\ref{subsubsec:heuristics}), the algorithm is able to be trained only with transitions sampled from the demonstrated expert policy. A specific instantiation of CSI (proposed Sec.~\ref{subsec:instanciation}) is tested on an inverted pendulum simulator (Sec.~\ref{subsec:IP}) and a car driving simulator (Sec.~\ref{subsec:Highway}) where we compare it with an AL algorithm \cite{abbeel2004apprenticeship}, with a pure classification algorithm \cite{taskar2005learning} and with a recent successfull IRL method \cite{klein2012scirl} as well as with a random baseline.

@@

We recall that solving the MDP is the direct problem. This contribution aims at solving the inverse one. We observe an expert's deterministic\footnotemark[\value{footnote}] policy $\pi_E$, assuming that there exists some unknown reward $R_E$ for which the expert is optimal. The suboptimality of the expert is an interesting setting that has been discussed for example in \cite{melo2010analysis,syed2010reduction}, but that we are not addressing here. We do not try to find this unknown reward $R_E$ but rather a non trivial reward $R$ for which the expert is at least near-optimal. The trivial reward $0$ is a solution to this ill-posed problem (no reward means that every behavior is optimal). Because of its ill-posed nature, this expression of \emph{Inverse Reinforcement Learning} (IRL) still has to find a satisfactory solution although lots of progress have been made, see Sec.~\ref{sec:related}.
\section{The cascading algorithm}
\label{sec:algo}
Our first step towards a reward function solving the IRL problem is a classification step using a \emph{score function-based multi-class classifier} (SFMC$^2$ for short). This classifier learns a score function $q\in\mathbb{R}^{S\times A}$ that rates the association of a given label $a\in A$ with a certain input $s\in S$. The classification rule $\pi_C\in A^S$ simply selects (one of) the label(s) that achieves the highest score for the given inputs:
\begin{equation}
  \label{eq:greedy2}
\pi_C(s) \in \argmax_a q(s,a).
\end{equation}
For example, \emph{Multi-class Support Vector Machines}~\cite{guermeur2011generic} can be seen as SFMC$^2$ algorithms, the same can be said of the structured margin approach \cite{taskar2005learning} which we consider in the experimental setting. Other algorithms may be envisionned (see Sec.~\ref{subsec:instanciation}).

@@

The only constraints the bound of Thm.~\ref{thm} implies on datasets $D_R$ and $D_C$ is that they provide enough information to the supervised algorithms to keep both error terms $\epsilon_C$ and $\epsilon_R$ low. In Sec.~\ref{sec:experiments} we deal with a dire lack of data in dataset $D_R$. We address the problem with the use of heuristics (Sec.~\ref{subsubsec:heuristics}) in order to show the behavior of the CSI algorithm in somewhat realistic conditions.

@@

%% Feel good comment
\section{Experiments}
\label{sec:experiments}
In this section, we put our approach under stress to empirically demonstrate its behavior. We begin by providing information pertaining to both benchmarks: an explanation about the amount and source of the available data is given on Sec.~\ref{subsubsec:data} and a precise description of the way we instanciated the CSI algorithm is given on Sec.~\ref{subsubsec:subroutines} and \ref{subsubsec:heuristics}.

We then use a classical RL benchmark, the Inverted Pendulum (Sec.~\ref{subsec:IP}), to provide a qualitative illustration of the proposed CSI approach.

We finally provide quantitative results and comparison with state-of-the-art approaches on a Highway driving simulator (Sec.~\ref{subsec:highway}).
\subsection{Instanciation}
\label{subsec:instanciation}
\subsubsection{Data scarcity}
\label{subsubsec:data}
In order to create a somewhat realistic environment, we decided that the CSI algorithm could neither use the transition probabilities nor a simulator. The only data it could access is sampled according to the expert policy. This corresponds for example to a situation where a costly system can only be controlled by a trained operator as a bad control sequence could lead to a system breakage.

@@

\subsection{Inverted Pendulum}
\label{subsec:IP}
We provide here qualitative results on the inverted pendulum problem. Quantitative results obtained by following a procedure very similar to Sec.~\ref{subsec:highway} were of the same nature and justified the same analysis as those provided in Sec.~\ref{subsubsec:hresults}. We thus omit them here for the sake of avoiding redundancy. Instead, we use the inverted pendulum problem for illustrative purposes.
\subsubsection{Setting}
The inverted pendulum (also known as the cart pole) is a classical benchmark in the RL community, it is fully described in \cite{lagoudakis2003least}. The goal is to maintain a pole in a unstable equilibrium position with the help of a cart moving along one axis. The state space consists in the angular position an speed of the pole. It is possible to use an RL algorithm to learn how to balance the pendule in an upright position by using the following reward: as long as the angular position of the pole is (in absolute value) less than $\pi \over 2$ the reward is $0$, the reward is $-1$ if the pole falls ; this reward function is plotted on Fig.~\ref{toto}. Three actions are available to the agent: it can choose to apply a leftward or rightward force on the cart or no force at all.

@@

\subsection{Highway driving simulator}
\label{subsec:highway}
Let's now move from a qualitative illustration to a quantitative assessment of the performance of the CSI algorithm. The Highway driving problem is described Sec.~\ref{subsubsec:hsetting}, the criterion we use is explained Sec.~\ref{subsubsec:hcriterion} and finally results and comparison to other approaches are presented Sec.~\ref{subsubsec:hresults}.
\subsubsection{Setting}
\label{subsubsec:hsetting}
The setting of the experiment is a driving simulator inspired from a benchmark already used in \cite{syed2008apprenticeship,syed2008game}. The agent controls a car that can switch between the three lanes of the road, go off-road on either side and modulate between three speed levels. At all times, there will be one car in one of the three lanes. Even at the lowest speed, the player's car moves faster than the others. When the other car disappears at the bottom of the screen, another one appear at the top in a randomly chosen lane. It takes two transitions to completely change lanes, as the player can move left or right for half a lane's length at a time. At the highest speed setting, if the other car appear in the lane the player is in, it is not possible to avoid the collision. The main difference between the original benchmark \cite{syed2008apprenticeship,syed2008game} and ours is that we made the MDP more ergodic by allowing the player to change speed whenever he wishes so, not just during the first transition. If anything, by adding two actions, we enlarged the state-action space and thus made the problem harder.
%FIXME:METTRE UNE CAPTURE DECRAN.
The reward function $R_E$ trains the expert to go as fast as possible (high reward) while avoiding collisions (harshly penalised) and avoiding going off-road (moderately penalised). Any other situation receives a null reward.

@@

  \caption{Data is shown with mean, standard deviation, minimum and maximum value over 50 runs. See Sec.~\ref{subsec:setting} for detailled information about the experimental setting.}
  \label{fig:Highway}
\end{figure}

@@

Classification and IRL have met in the past in \cite{ratliff2006maximum}, but the labels were complete optimal policies rather than actions and the inputs were MDPs, which had to be solved. The SCIRL (for Structured Classification for IRL) algorithm \cite{klein2012scirl} which we used Sec.~\ref{subsec:highway} is able to solve the IRL problem (without having to solve any MDP) in the same dire conditions the cascading approach has been tested in. It injects the structure of the MDP in a linearly parametrized score-function based classifier by using the expert's feature expectation as feature function. This necessity of a linarly parametrized approach and the need to compute the expert's feature exepctations makes the SCIRL algorithm less convenient than the CSI approach.
